Global patterns of forest autotrophic carbon fluxes

### 2020-03-02 manuscript/manuscript draft.Rmd
edited results (finished)

(*KAT has revised results up to here.*)

There was a significant interactive effect between MAT and MAP for NPP and $ANPP_{woody-stem}$ (fig.3). There was a significant additive effect for GPP, ANPP and autotrophic respiration. $ANPP_{foliage}$, BNPP, $BNPP_{root-fine}$, and root respiration showed no significant interactive or additive effect (see table S3).

For the variables which showed a significant interactive or additive effect between MAT and MAP, no other climate variable, in combination with MAT, significantly improved on that model. {**need to confirm this given changes in MAT MAP results**}

@@

*How do FACF relate to other climate variables?* {**Here, maybe add some statement about covariation among climate variables, site SI figure**} Our results indicated that FACF were most strongly explained by temperature at the global scale, with temperature-related climate variables coming out as strong predictors of FACF. In addition to MAT, temperature seasonality, annual temperature range, and annual frost days were consistently identified as strong univariate predictors of FACF.  {**cite some table/ SI?**}

We found a significant relationship between C flux and potential evapotranspiration for all FACF. $ANPP_{foliage}$, $BNPP_{root-fine}$ and $R_{root}$ increased linearly with PET; however, all other fluxes showed a polynomial relationship with PET (fig. 4). We found strong evidence for a saturation point or peak with PET: FACF tended to increase at values below 1000mm, before saturating between 1200 and 1700mm. There was also evidence that FACF begin to decrease at values above 1800mm PET.

Vapour pressure deficit was a significant predictor of C flux for all FACF. $BNPP_{root-fine}$ showed a linear relationship with vapour pressure deficit ($R^2$ = 0.07, p<0.05), but all other fluxes showed a polynomial relationship (fig. 4). FACF initially increased with vapour pressure deficit, before saturating at around 0.8 kPa. At values above 0.8 kPa, FACF began to decrease.

All fluxes, with the exception of $R_{root}$, showed a positive linear relationship with solar radiation. Solar radiation explained a low proportion of variability in all FACF, explaining less than 20% of the variation in each flux, with the exception of$R_{auto}$ ($R^2$ = 0.26, p<0.05).

Of the climate variables tested, annual wet days, aridity, cloud cover, precipitation seasonality, maximum vapour pressure deficit and water stress months were poor or non-significant explainers of variation in FACF, explaining less than 20% of the variation in each of the carbon fluxes.

@@

*What is the role of seasonality in explaining FACF?* Temperature seasonality was a significant predictor of FACF. GPP, NPP, ANPP, and $R_{root}$ exhibited a polynomial relationship with seasonality (Fig. 4) {**give statistic on signficiance**}. $ANPP_{foliage}$, $ANPP_{woody-stem}$ and $R_{auto}$ decreased linearly with temperature seasonality. Temperature seasonality was strongly correlated with annual temperature range, and, as expected, all fluxes showed almost identical responses to it {**meaning unclear**}. FACF were highest where temperature seasonality = 0, and at an annual temperature range of 15$^\circ$C or lower. In contrast, there was no significant effect of precipitation seasonality on FACF.

We found a significant relationship between length of growing season and FACF, with all fluxes showing a linear increase with length of growing season (Fig. 4). Length of growing season was a strong predictor of FACF, explaining 51% of variation in GPP, 39% of variation in NPP, and 34% of variation in ANPP, but it was a weaker predictor than MAT for all fluxes analysed.

*Within the growing season, how do FACF vary with climate?* When FACF were standardized by growing season length, correlations with climate variables were much weaker. Growing season temperature, precipitation, solar radiation, and PET each had a small effect on one or more of the nine FACF analyzed (Table 2; Fig. **S#**). Speficifally, $ANPP$ increased with growing season temperature ($R^2$ = 0.10, p<0.001) and precipitation ($R^2$ = 0.04, p<0.05). Similaryly, $ANPP_{foliage}$ increased slightly with growing season temperature ($R^2$ = 0.16, p<0.01) and precipitation ($R^2$ = 0.09, p<0.05). Growing season solar radiation had a positive influence on $GPP$ ($R^2$ = 0.21, p<0.001), $NPP$ ($R^2$ = 0.21, p<0.001), $BNPP$ ($R^2$ = 0.16, p<0.001) and $BNPP_{fine.root}$ ($R^2$ = 0.12, p<0.01). Growing season PET had a positive influence on $GPP$ ($R^2$ = 0.15, p<0.01), $NPP$ ($R^2$ = 0.18, p<0.01), $BNPP$ ($R^2$ = 0.23, p<0.0001), $BNPP_{fine.root}$ ($R^2$ = 0.11, p<0.05), and $ANPP_{woody-stem}$ ($R^2$ = 0.06, p<0.05). {**Becky, please verify/ edit the following:** There were no other signifcant correlations between growing season length-standardized FACF (9 variables in Table 2) and growing season climate (**which variables?**)}.

>>>>

There was a significant additive effect of MAT and MAP on $GPP$, $ANPP$ and $R_{auto}$ (Fig. 3, Table S3). Accounting for MAT, MAP had a substantial positive effect on $GPP$ and $R_{auto}$ and a small negative effect on $ANPP$. There was a significant interactive effect between MAT and MAP for $NPP$ and $ANPP_{woody-stem}$ (Fig. 3, Table S3). The interaction was negative for $NPP$ and positive for $ANPP_{woody-stem}$. For $ANPP_{foliage}$, $BNPP$, $BNPP_{root-fine}$, and $R_{auto-root}$, MAP did not have a signficant effect when accounting for MAT (Fig. 3, Table S3). For the variables which showed a significant interactive or additive effect between MAT and MAP, no other climate variable, in combination with MAT, significantly improved on that model. {**need to confirm this given changes in MAT MAP results (or you could just drop the sentence.)**}

@@

*How do FACF relate to other climate variables?*

Our results indicated that FACF were most strongly explained by temperature at the global scale, with temperature-related climate variables coming out as strong predictors of FACF. In addition to MAT, several of its correlates (Fig. S2) were consistently identified as strong univariate predictors of FACF: temperature seasonality, annual temperature range, annual frost days, PET, and length of growing season (Figs. 4, S4-S7).

We found a significant relationship between C flux and potential evapotranspiration for all FACF. $ANPP_{foliage}$, $BNPP_{root-fine}$ and $R_{root}$ increased linearly with PET; however, all other fluxes showed a polynomial relationship with PET (Fig. 4c, S4-5; Table S2). We found strong evidence for a saturation point or peak with PET: FACF tended to increase at values below 1000mm, before saturating between 1200 and 1700mm. There was also evidence that FACF begin to decrease at values above 1800mm PET.

Vapour pressure deficit was a significant predictor of C flux for all FACF. $BNPP_{root-fine}$ showed a linear relationship with vapour pressure deficit ($R^2$ = 0.07, p<0.05), but all other fluxes showed a polynomial relationship (Figs. 4d, S4-5; Table S2). FACF initially increased with vapour pressure deficit, before saturating at around 0.8 kPa, after wich point they began to decrease.

All fluxes, with the exception of $R_{root}$, showed a positive linear relationship with solar radiation (Figs. S4-S5, Table S2). Solar radiation explained a low proportion of variability in all FACF, explaining less than 20% of the variation in each flux, with the exception of$R_{auto}$ ($R^2$ = 0.26, p<0.05).

Aannual wet days, cloud cover, and aridity were poor or non-significant explainers of variation in FACF, explaining less than 20% of the variation in each of the carbon fluxes (Figs. S4-S5; Table S2).

@@

*What is the role of seasonality in explaining FACF?*

Temperature seasonality was a significant predictor of FACF. $GPP$, $NPP$, $ANPP$, and $R_{root}$ exhibited a polynomial relationship with seasonality (all p<0.05; Figs. 4e, S6-7; Table S2). $ANPP_{foliage}$, $ANPP_{woody-stem}$ and $R_{auto}$ decreased linearly with temperature seasonality (all p<0.05; Figs. 4e, S6-S7; Table S2). Temperature seasonality was strongly correlated with annual temperature range, which was likewise a similarly strong predictor of FACF (Table S2).FACF were highest where temperature seasonality = 0, and at an annual temperature range of 15$^\circ$C or lower.

In contrast, there was no significant effect of precipitation seasonality on FACF, and both maximum vapour pressure deficit, and water stress months were poor or non-significant explainers of variation in FACF (Figs. S6-S7; Table S2).

We found a significant relationship between length of growing season (temperature-defined) and FACF, with all fluxes showing a linear increase with length of growing season (Figs. 4e, S6-S7; Table S2). Length of growing season was a strong predictor of FACF, explaining 51% of variation in GPP, 39% of variation in NPP, and 34% of variation in ANPP, but it was a weaker predictor than MAT for all fluxes analysed.

*Within the growing season, how do FACF vary with climate?*

When FACF were standardized by growing season length, correlations with growing season climate--includingtemperature, precipitation, solar radiation, and PET--were generally weak (Table S4, Figs. S8-S9). Speficifally, $ANPP$ increased with growing season temperature ($R^2$ = 0.10, p<0.001) and precipitation ($R^2$ = 0.04, p<0.05). Similaryly, $ANPP_{foliage}$ increased slightly with growing season temperature ($R^2$ = 0.16, p<0.01) and precipitation ($R^2$ = 0.09, p<0.05). Growing season solar radiation had a positive influence on $GPP$ ($R^2$ = 0.21, p<0.001), $NPP$ ($R^2$ = 0.21, p<0.001), $BNPP$ ($R^2$ = 0.16, p<0.001) and $BNPP_{fine.root}$ ($R^2$ = 0.12, p<0.01). Growing season PET had a positive influence on $GPP$ ($R^2$ = 0.15, p<0.01), $NPP$ ($R^2$ = 0.18, p<0.01), $BNPP$ ($R^2$ = 0.23, p<0.0001), $BNPP_{fine.root}$ ($R^2$ = 0.11, p<0.05), and $ANPP_{woody-stem}$ ($R^2$ = 0.06, p<0.05). {**Becky, please verify/ edit the following:** There were no other signifcant correlations between growing season length-standardized FACF (9 variables in Table 2) and growing season climate (**which variables?**)}.
